---
title: "Analysis for test run data"
author: "Enuh"
date: "October 1, 2021"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####Preparatory####


**1.Preparatory**
```{r}
##Load requirements
library(readxl)
datapath = "C:/Users/Blaisonyl/Desktop/Help/David_Research/DataAndReports/Data_set_trap_color_and_bait.xlsx"
data <- read_excel(datapath)
head(data)
View(data)
```




```{r}
library(dplyr)
library(ggplot2)
library(mice)
library(tidyverse)
library(janitor)
library(kableExtra)
library(AER)
library(vegan)
library(fossil)
library(tinytex)
```

```{r}
##check missing data
missing = is.na(data)
missing2 = data %>% filter(!complete.cases(data))
```

#Give the column names the appropriate syntax names temporarily store in dew df called data_new1
```{r}
data_new1 <- data                  
colnames(data_new1) <- gsub(" ", "_", colnames(data_new1))
head(data_new1)                       
data2 = data_new1
```

#Convert chars to factor
```{r}
data2[sapply(data2, is.character)] <- lapply(data2[sapply(data2, is.character)], as.factor)

#########Summary of data i need 
data2 %>% tabyl(species_id, Trap_type)
data2 %>% tabyl(species_id, Trap_colour)
data2 %>% tabyl(species_id, sex)

###'<U+2640>' = 'female'
###'<U+2642>' = 'male'
```


```{r}
#datset to males and females
Females = data2 %>% split(data2$sex_coded)
FemalesData = Females$`1`
MalesData = Females$`2`

```



####Trap types####

**2. Trap_Type**

**2.1 Abundance of species per Trap type**
```{r}
data2 %>% tabyl(species_id, Trap_type) %>% {. ->> Trap_type } %>% 
  adorn_totals( where = 'row') %>%
  adorn_totals( where = 'col') %>%
  kbl(caption = "Abundance of species per Trap_type") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```
```{r}
Trap_type %>%
  rowwise() %>% 
  mutate(
    X2 = chisq.test(c(`net trap`, `resting cage`))$statistic,
    p_value = chisq.test(c(`net trap`, `resting cage`))$p.value
    ) %>% kbl(caption = "Trap type") %>% kable_classic(full_width = F, html_font = "Times new roman")

```



```{r}  
#Preprocessing and ANOVA
colnames(Trap_type)<- c('species_id','net_trap','resting_cage')
a = gather(Trap_type, "Traps", "Abundance", !1)
oneway <- aov(Abundance ~ Traps, data = a)
summary(oneway)
```


```{r} 
#General linear model for Abundance and trap type
Trap_Glm = glm(Abundance ~ Traps, data = a)
Trap_Glm
```


**2.2 Diversity according to Trap type**
```{r}
#Shannon index
data2 %>% tabyl(Trap_type, species_id) %>% 
  untabyl() %>% 
  select(!Trap_type)%>% 
  diversity(index = "shannon") %>% 
  cbind(c('net_trap', 'resting_cage' )) %>%  
  kbl(caption = "Shannon indexes per Trap type") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```

```{r}
#Shannon index
data2 %>% tabyl(Trap_type, species_id) %>%   
  untabyl() %>% 
  select(!Trap_type)%>% 
  diversity(index = "shannon") %>% 
  cbind(c('net_trap', 'resting_cage' )) %>%  
  kbl(caption = "Shannon indexes per Trap type") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```


```{r}
#Simpson index
data2 %>% tabyl(Trap_type, species_id) %>% 
  untabyl() %>% 
  select(!Trap_type)%>% 
  diversity(index = "simpson") %>% 
  cbind(c('net_trap', 'resting_cage')) %>%  
  kbl(caption = "Simpson indexes per Trap type") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```


```{r}
#ACE 
data2 %>% tabyl(species_id, Trap_type) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, ACE)
```

```{r}
#Chao1
data2 %>% tabyl(species_id, Trap_type) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, chao1)
```

***Trap type Diversity females***
```{r}
#Shannon index
FemalesData %>% tabyl(Trap_type, species_id) %>%   
  untabyl() %>% 
  select(!Trap_type)%>% 
  diversity(index = "shannon") %>% 
  cbind(c('net_trap', 'resting_cage' )) %>%  
  kbl(caption = "Shannon indexes per Trap type") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```

```{r}
#Simpson index
FemalesData %>% tabyl(Trap_type, species_id) %>% 
  untabyl() %>% 
  select(!Trap_type)%>% 
  diversity(index = "simpson") %>% 
  cbind(c('net_trap', 'resting_cage')) %>%  
  kbl(caption = "Simpson indexes per Trap type") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```


```{r}
#ACE 
FemalesData %>% tabyl(species_id, Trap_type) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, ACE)
```

```{r}
#Chao1
FemalesData %>% tabyl(species_id, Trap_type) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, chao1)
```


***Trap type Diversity Males***
```{r}
#Shannon index
MalesData %>% tabyl(Trap_type, species_id) %>%   
  untabyl() %>% 
  select(!Trap_type)%>% 
  diversity(index = "shannon") %>% 
  cbind(c('net_trap', 'resting_cage' )) %>%  
  kbl(caption = "Shannon indexes per Trap type") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```


```{r}
#Simpson index
MalesData %>% tabyl(Trap_type, species_id) %>% 
  untabyl() %>% 
  select(!Trap_type)%>% 
  diversity(index = "simpson") %>% 
  cbind(c('net_trap', 'resting_cage')) %>%  
  kbl(caption = "Simpson indexes per Trap type") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```


```{r}
#ACE 
MalesData %>% tabyl(species_id, Trap_type) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, ACE)
```

```{r}
#Chao1
MalesData %>% tabyl(species_id, Trap_type) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, chao1)
```





####trap colour####

**3. Trap colour**

**3.1 Abundance of species per Trap colour**
```{r}
data2 %>% tabyl(species_id, Trap_colour) %>% {. ->> Trap_colour } %>% 
  adorn_totals( where = 'row') %>%
  adorn_totals( where = 'col') %>%
  kbl(caption = "Abundance of species per Trap_colour") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```
```{r}

Trap_colour %>%
  rowwise() %>% 
  mutate(
    X2 = chisq.test(c(`black`,`blue`,`green`,`pink`,`white`))$statistic,
    p_value = chisq.test(c(`black`,`blue`,`green`,`pink`,`white`))$p.value
    ) %>% kbl(caption = "Trap colour") %>% kable_classic(full_width = F, html_font = "Times new roman")
```



```{r}  
#Preprocessing and ANOVA
colnames(Trap_colour)<- c("species_id","black","blue","green","pink","white")
a = gather(Trap_colour, "colours", "Abundance", !1)
oneway <- aov(Abundance ~ colours, data = a)
summary(oneway)
```

```{r} 
#General linear model for Abundance and trap colour
Trapcol_Glm = glm(Abundance ~ colours, data = a)
Trapcol_Glm
```


**3.2 Diversity according to Trap colour**
```{r}
#Shannon index
data2 %>% tabyl(Trap_colour, species_id) %>% 
  untabyl() %>% 
  select(!Trap_colour)%>% 
  diversity(index = "shannon") %>% 
  cbind(c("black","blue","green","pink","white")) %>%  
  kbl(caption = "Shannon indexes per Trap colour") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```


```{r}
#Simpson index
data2 %>% tabyl(Trap_colour, species_id) %>% 
  untabyl() %>% 
  select(!Trap_colour)%>% 
  diversity(index = "simpson") %>% 
  cbind(c("black","blue","green","pink","white")) %>%  
  kbl(caption = "Simpson indexes per Trap colour") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```

```{r}
#ACE 
data2 %>% tabyl(species_id, Trap_colour) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, ACE)
```

```{r}
#Chao1
data2 %>% tabyl(species_id, Trap_colour) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, chao1)
```

***Trap colour Diversity of females***
```{r}
#Shannon index
FemalesData %>% tabyl(Trap_colour, species_id) %>% 
  untabyl() %>% 
  select(!Trap_colour)%>% 
  diversity(index = "shannon") %>% 
  cbind(c("black","blue","green","pink","white")) %>%  
  kbl(caption = "Shannon indexes per Trap colour") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```


```{r}
#Simpson index
FemalesData %>% tabyl(Trap_colour, species_id) %>% 
  untabyl() %>% 
  select(!Trap_colour)%>% 
  diversity(index = "simpson") %>% 
  cbind(c("black","blue","green","pink","white")) %>%  
  kbl(caption = "Simpson indexes per Trap colour") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```

```{r}
#ACE 
FemalesData %>% tabyl(species_id, Trap_colour) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, ACE)
```

```{r}
#Chao1
FemalesData %>% tabyl(species_id, Trap_colour) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, chao1)
```


***Trap colour Diversity of Males***
```{r}
#Shannon index
MalesData %>% tabyl(Trap_colour, species_id) %>% 
  untabyl() %>% 
  select(!Trap_colour)%>% 
  diversity(index = "shannon") %>% 
  cbind(c("black","blue","green","pink","white")) %>%  
  kbl(caption = "Shannon indexes per Trap colour") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```


```{r}
#Simpson index
MalesData %>% tabyl(Trap_colour, species_id) %>% 
  untabyl() %>% 
  select(!Trap_colour)%>% 
  diversity(index = "simpson") %>% 
  cbind(c("black","blue","green","pink","white")) %>%  
  kbl(caption = "Simpson indexes per Trap colour") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```

```{r}
#ACE 
MalesData %>% tabyl(species_id, Trap_colour) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, ACE)
```

```{r}
#Chao1
MalesData %>% tabyl(species_id, Trap_colour) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, chao1)
```

####Baits####

**3. Baits**
  
**3.1 Diversity according to baits used**

```{r}
data2 %>% tabyl(species_id, baits_used) %>% {. ->> Bait } %>%
  adorn_totals( where = 'row') %>%
  adorn_totals( where = 'col') %>%
  kbl(caption = "Abundance of species per bait used") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```
```{r}
Bait %>%
  rowwise() %>% 
  mutate(
    X2 = chisq.test(c(`Carbon dioxide`,`chicken`,`no bait`,`pigeon`))$statistic,
    p_value = chisq.test(c(`Carbon dioxide`,`chicken`,`no bait`,`pigeon`))$p.value
    ) %>% kbl(caption = "Bait") %>% kable_classic(full_width = F, html_font = "Times new roman")
```



```{r}
#Preprocessing and ANOVA
Bait1 = Bait %>% gather( "Bait", "Abundance", !1) 
aov(Abundance ~ Bait, data = Bait1) %>% 
  summary()
#plot(Bait1$Bait ~ Bait1$Abundance, ylim = 0)
```

```{r} 
#General linear model for Abundance and bait type
Bait_Glm = glm(Abundance ~ Bait, data = Bait1)
Bait_Glm
```


**3.2 Diversity according to Bait used**

```{r}
#Shannon index
data2 %>% tabyl(baits_used, species_id) %>% 
  untabyl() %>% 
  select(!baits_used)%>% 
  diversity(index = "shannon") %>% 
  cbind(c("Carbon dioxide", "chicken","no bait","pigeon")) %>%  
  kbl(caption = "Shannon indexes per bait used") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```

```{r}
#Simpson index
data2 %>% tabyl(baits_used, species_id) %>% 
  untabyl() %>% 
  select(!baits_used)%>% 
  diversity(index = "simpson") %>% 
  cbind(c("Carbon dioxide", "chicken","no bait","pigeon")) %>%  
  kbl(caption = "Simpson indexes per bait used") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```

```{r}
#ACE 
data2 %>% tabyl(species_id, baits_used) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, ACE)
```

```{r}
#Chao1
data2 %>% tabyl(species_id, baits_used) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, chao1)
```

***bait use Diversity of females ***
```{r}
#Shannon index
FemalesData %>% tabyl(baits_used, species_id) %>% 
  untabyl() %>% 
  select(!baits_used)%>% 
  diversity(index = "shannon") %>% 
  cbind(c("Carbon dioxide", "chicken","no bait","pigeon")) %>%  
  kbl(caption = "Shannon indexes per bait used") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```

```{r}
#Simpson index
FemalesData %>% tabyl(baits_used, species_id) %>% 
  untabyl() %>% 
  select(!baits_used)%>% 
  diversity(index = "simpson") %>% 
  cbind(c("Carbon dioxide", "chicken","no bait","pigeon")) %>%  
  kbl(caption = "Simpson indexes per bait used") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```

```{r}
#ACE 
FemalesData %>% tabyl(species_id, baits_used) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, ACE)
```

```{r}
#Chao1
FemalesData %>% tabyl(species_id, baits_used) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, chao1)
```

***bait use Diversity of Males ***
```{r}
#Shannon index
MalesData %>% tabyl(baits_used, species_id) %>% 
  untabyl() %>% 
  select(!baits_used)%>% 
  diversity(index = "shannon") %>% 
  cbind(c("Carbon dioxide", "chicken","no bait","pigeon")) %>%  
  kbl(caption = "Shannon indexes per bait used") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```

```{r}
#Simpson index
MalesData %>% tabyl(baits_used, species_id) %>% 
  untabyl() %>% 
  select(!baits_used)%>% 
  diversity(index = "simpson") %>% 
  cbind(c("Carbon dioxide", "chicken","no bait","pigeon")) %>%  
  kbl(caption = "Simpson indexes per bait used") %>%
  kable_classic(full_width = F, html_font = "Times new roman")
```

```{r}
#ACE 
MalesData %>% tabyl(species_id, baits_used) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, ACE)
```

```{r}
#Chao1
MalesData %>% tabyl(species_id, baits_used) %>% 
  untabyl() %>% select(!species_id)%>% apply( MARGIN = 2, chao1)
```
